
\section{The problems}\label{sec:problems}

In this section we summarize the main problems identified in CPDOC's
current infrastructure and daily working environment.

As described in Section~\ref{sec:cpdoc}, CPDOC's archives are
maintained in three different information systems based on traditional
relational data models. This infrastructure is hard to maintain,
improve and refine. 
% The information systems do not provide an intuitive and effective
% interfaces to their users, making the access to information difficult
% and limited.

% The information systems were designed and developed in the 90's and
% since then has undergone several modifications. The changes were
% motivated by different reasons: from requirements of users to
% substitution of obsolete technologies (database systems or web
% development frameworks and programming languages). A careful
% analysis of CPDOC systems would consider them to be outdated
% especially if compared with the current available systems and
% technologies for data storage, indexing and long-term preservation.

The data of CPDOC's collections  is
actually limited to what is called ``Deep
Web''~\cite{bergman2001white}. The information cannot be found or
access by standard search engines for mainly two reasons: (1) the page
does not exist until being created dynamically as a result of an
specific query; (2) the users are required to login in order to make
queries or access the digital files. Service providers does not access
the data directly and therefore cannot provide specialized services
using it. Users themselves are not able to expand the queries
over the collections, being limited to the avaiable interfaces in the website.

Concerning terminology, CPDOC's collections do not follow any metadata
standards, which hinder considerably the interoperability with other
digital sources.

Queries in general face idiosyncratic indexing
problems and low rates of recall and precision. These problems are
basically linked to the \emph{ad hoc} indexing strategy adopted
earlier to define database tables and fields.

% The maintenance of 3 different information systems is expensive,
% time demanding and ineffective. Improvements are hard to implement
% and therefore innovative initiatives are usually postponed. We
% believe this is a side effect of the adoption of a relational data
% model and the traditional information systems maintenance
% methodology.
% % Regarding the items above, it is perhaps worth justifying in
% % technical detail the last one.
% In a relational database system, it is well known that it is
% expensive and hard to frenquently change the data model. This is due
% to the assumption that a relational data model must be defined
% \emph{a priori}, i.e., before the beginning of data
% acquisition. Moreover, information systems for maintenance of
% relational databases are also difficult to evolve. Changes in the
% database require adaptations in the system interfaces and
% reports. This workflow is expensive, time consuming and demands
% several different professionals with different skills: interface
% developers, database administrators etc.

The maintenance of 3 different information systems is expensive, time
demanding and ineffective. Improvements are hard to implement and
therefore innovative initiatives are usually postponed.
% Regarding the items above, it is perhaps worth justifying in
% technical detail the last one.
The truth is that in a relational database system, it is not easy to
freely change the data model, due to the
fact that a relational data model must be defined \emph{a
  priori}, i.e., before the  data acquisition's stage. Moreover,
changes in the database require adaptations in
the system interfaces and reports. The workflow is expensive, time
consuming and demands several different professionals with different
skills: interface developers, database administrators etc.

% The CPDOC researchers rely on the FGV's technical staff team from
% the IT department for any database modification or user interface
% improvement. Even the addition of a new field in a table requires a
% whole workflow comprising: (1) the specification of the requirement;
% (2) the costs and time estimation from the technical staff; (3) the
% appropriate authorization; (4) the allocation of a technical
% profissional or team for a period of time considering the other
% FGV's priorities and demands. This workflow usually takes weeks and
% even months.
%
% Whenever distributed and interoperable data maintenance is required,
% the relational data model fails severely. First, it is not easy to
% keep the model and data independent. Second, data exchange is
% hindered by then different storage strategies for datatypes and
% encodes of strings. Third, relational data models are commonly
% flooded with several auxiliary tables necessary to store non-trivial
% relations like N-N relations or generalizations. And finally, in
% general, few model constrains are maintained in databases, and the
% ones created are also not easily interoperable.

% Concerning data storage, it is also worth mentioning that Acessus
% and Oral Historic Interviews collections are not stored in a single
% place. The actual files (digitalized documents) of personal archives
% and the files of digitalized interviews are scattered in different
% file servers.

Concerning data storage, it is also worth mentioning that the
digitalized Acessus's documents and the Oral History's interviews are
not stored in a single place, but scattered in two different file
servers. The CPDOC database only stores the metadata and paths in the
file servers, which hinders the maintenance of consistency between
files, metadata information and access control policies.

% Whereas people make great use of standard search engines to query
% and share knowledge, it becomes certainly one of the greatest
% challenges faced by CPDOC when it comes to open their collections.

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "article"
%%% End: 
